---
title: "ACM_Beginning_DS_Project"
author: "Khushi Chaudhari"
date: "2025-02-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Imports

```{r}
# uncomment the 2 lines below and run ONCE

install.packages("ggplot2")
install.packages("tidyverse")

# great package for visualizations
library(ggplot2)
# great package for data manipulation
library(tidyverse)
```
ggplot2 -> the package to visualize the data


```{r}
# imports the dataset into variable "df"
# update the path to match path of your local file

df_uncleaned <- read.csv("ACM Data Science Survey (Responses) - Form Responses 1.csv")
colnames(df_uncleaned) <- c("time_stamp", "identity", "distance", "transportation", 
                  "parking_satisfication", "activity_participation", "org_count", 
                  "social_satisfication", "sleep_amount", "meals_count")
```

```{r}
# check how many data has empty string/NA
sum(df_uncleaned == "", na.rm = TRUE)
# convert "" into NA
df_uncleaned[df_uncleaned == ""] <- NA
# check again
sum(df_uncleaned == "", na.rm = TRUE)
sum(is.na(df_uncleaned))  # answer: 4

# delete the rows that has NA included
df <- na.omit(df_uncleaned)

# check again, should be 0 for both NA and ""
sum (is.na(df)) 
sum(df == "", na.rm = TRUE)

df
```

## Hypothesis: Living Situation and Social Fulfillment
Students living on-campus report feeling more socially fulfilled (higher ratings on a scale of 1-5) 
and being more involved to the on-campus organizations compared to commuters.

Null Hypothesis (H₀):
There is no significant difference in social satisfaction and involvement in 
campus organizations between studnets who live on-campus or commute.

variables will be using:

1. identity (chr): commuter vs on-campus
2. activity_participation (char): yes/no/maybe
3. social_satisfication (int): scale 1-5 about social life satification
4. org_count(chr) : 0-1/2-3/...


```{r}
# 1. identity vs activity_participation, 2 categorical -> bar graph
ggplot(data = df, aes(x=identity, fill=activity_participation)) + 
  geom_bar(position="dodge") +  # Use "stack" for stacked bars
  labs(title="Living Situation vs Activity Participation", x="Living Situation", y="Activity Participation")

```

```{r}
# 2. identity vs social_satisfication, 1 numerical 1 categorical -> bot plot/violin plot
ggplot(df, aes(x=identity, y=social_satisfication)) +
  geom_dotplot(binaxis="y", stackdir="center", dotsize=0.5) +
  labs(title="Living Situation vs Overall Participation (dot plot)", x="Living Situation", y="Social Satisfaction")

ggplot(df, aes(x=identity, y=social_satisfication, fill=identity)) +
  geom_violin(alpha=0.5) +
  labs(title="Living Situation vs Overall Participation (violin plot)", x="Living Situation", y="Social Satisfaction")

```
```{r}
# 3. identity vs org_count, 2 categorical -> bar graph
ggplot(data = df, aes(x=identity, fill=org_count)) + 
  geom_bar(position="dodge") + 
  labs(title="Living Situation vs Organization Count", x="Living Situation", y="Organization Count")
```

```{r}
# 4. overall_participation: activity_participation * org_count(chr), 1 numerical 1 categorical -> bot plot/violin plot
# convert the activity_participation into numeric variable: yes = 2, no = 1, maybe = 1.5
# convert the org_count into numeric variable: average of the interval

activity_vector <- ifelse(df$activity_participation == "Yes", 2, 
                          ifelse(df$activity_participation == "No", 1, 
                          1.5))

org_count_numeric <- ifelse(df$org_count == "0-1", 0.5, 
                            ifelse(df$org_count == "2-3", 2.5, 
                                   ifelse(df$org_count == "4-5", 4.5, 
                                          6)))

overall_participation <- activity_vector * org_count_numeric

ggplot(df, aes(x=identity, y=overall_participation)) +
  geom_dotplot(binaxis="y", stackdir="center", dotsize=0.5) +
  labs(title="Living Situation vs Overall Participation (dot plot)", x="Living Situation", y="Overall Participation")

ggplot(df, aes(x=identity, y=overall_participation, fill=identity)) +
  geom_violin(alpha=0.5) +
  labs(title="Living Situation vs Overall Participation (violin plot)", x="Living Situation", y="Overall Participation")


```


